+++
title = "✓ Reshape Data"
icon = "sticky-fill"
icon = "r-project"
tags = ["reshape", ]

link = "https://nbviewer.org/github/paradoxical-rhapsody/jupyterNBs/blob/master/211104-reshape.ipynb"
+++

# Reshape Data
*2021-11-04*

The repeated measurements in observed data can be orginazed as `wide` or `long` format:

* `wide`: the repeated measurements are arranged into sepearated columns of the same row.
* `long`: repeated measurements are arranged into sepearated rows of the same column.

Two ways to tranform the data between long and wide format are considered:

1. Built-in fuction `reshape` is qualified to the conversion.
2. Pkg `data.table` provide function `melt` (wide-to-long) and `dcast` (long-to-wide).


## `stats::reshape`

Let `DF` be a wide data, i.e.

```{r}
rm(list=ls())
set.seed(2021)
DF <- matrix(rnorm(3*9), 3) |> round(digits=2) |> data.frame()
colnames(DF) <- paste(rep(c('F', 'G', 'H'), each=3), c('a', 'b', 'c'), sep='.')

DF <- data.frame(ID=letters[1:3], DF)
show(DF)
```

In the cell below the `DF` is melted into a long table.

```{r}
varying <- list(
    c('F.a', 'G.a', 'H.a'),
    c('F.b', 'G.b', 'H.b'),
    c('F.c', 'G.c', 'H.c')
)

times <- c('F', 'G', 'H')

v.names <- c('a', 'b', 'c')
timevar <- 'method'

DF.long <- reshape(DF, direction='long', varying=varying, times=times, v.names=v.names, timevar=timevar)
show(DF.long)
```


## `data.table::melt/dcast`

```{r}
library(data.table)
DT <- copy(DF) |> as.data.table()

DT.long <- melt(DT, id.vars='ID', 
    measure.vars=list(
        a=c('F.a', 'G.a', 'H.a'), 
        b=c('F.b', 'G.b', 'H.b'), 
        c=c('F.c', 'G.c', 'H.c')), 
    variable.name='method')

DT.long[method==1, method:='F']
DT.long[method==2, method:='G']
DT.long[method==3, method:='H']

show(DT.long)
```

```{r}
show(dcast(DT.long, ID ~ method, value.var=c('a', 'b', 'c')))
```

